Towards Practical First-Order Model Counting

### 2025-07-05 doc/talk/talk.tex
Finish explaining the workflow

\documentclass{beamer}
\usepackage[utf8]{inputenc}
\usepackage[UKenglish]{babel}
\usepackage[UKenglish]{isodate}
\usepackage{tikz}
\usepackage[linesnumbered,ruled,vlined]{algorithm2e}
\usepackage{complexity}
\usepackage{fontawesome}

@@

@@

\begin{frame}{Knowledge Compilation Workflow}
  \centering
  \begin{tikzpicture}
    % Top row
    \node[anchor=west] at (-0.5, 0) (formula) {$\phi$};
    \node[draw,rounded rectangle] at (3, 0) (compilewithbasecases) {\CompileWithBaseCases};
    \node[draw,rounded rectangle] at (8, 0) (compilation) {Compile to C++};

    % Bottom row
    \node[draw,rounded rectangle,dashed] at (8, -3) (cpp) {C++ code};
    \node[anchor=west] at (-1, -3) (sizes) {Domain sizes};
    \node at (8, -4) (count) {Model count};

    % Modules
    \node[draw,rounded rectangle,anchor=north west] at (1.2, -1) (crane) {\Compile};
    \node[draw,rounded rectangle,right=0.1cm of crane] (findbasecases) {\FindBaseCases};
    \node[draw,rounded rectangle,anchor=north west] at (1.2, -1.6) (propagate) {\Propagate};
    \node[draw,rounded rectangle,right=0.1cm of propagate] (simplify) {\Simplify};

    % Bounding box and its name
    \begin{scope}[on background layer]
      \node[draw,fit={(compilewithbasecases) (compilation) (crane) (findbasecases) (propagate)},inner ysep=7pt,yshift=5pt,fill=gray!10] {};
    \end{scope}
    \node at (1.5, 0.5) {\Cranetwo};

    % Brace and its arrow
    \node[fit=(crane)(findbasecases)(propagate)(simplify)] (uses) {};
    \draw[decorate,decoration={brace}] (uses.north west) -- (uses.north east) node[midway] (brace) {};
    \draw[-Latex,dashed] (compilewithbasecases) -- node[midway,left] {uses} (brace);

    % All other arrows
    \draw[-Latex] (formula) -- (compilewithbasecases);
    \draw[-Latex] (compilewithbasecases) -- node[above] {$\mathcal{E}$} (compilation);
    \draw[-Latex] (compilation) -- (cpp);
    \draw[-Latex] (sizes) -- (cpp);
    \draw[-Latex] (cpp) -- (count);
  \end{tikzpicture}
\end{frame}

% TODO: 4 more slides explaining the solution (maybe including the theorems)
% 0. Preprocess
% 1. Use Crane to compile into functions
% 2. Simplify
% 3. FindBaseCases TODO
% 4. For each base case
% 4.1. For each constant (TODO?)
% 4.1.1. Propagate TODO
% 5. Compile to C++

\begin{frame}{Knowledge Compilation Workflow (1/2)}
  \begin{enumerate}
    \item Use \textsc{Crane} to compile \structure{$\phi$} into a set of
          equations \structure{$\mathcal{E}$}
    \item Simplify them, e.g.,
          \[
            g(l, m) = \sum_{k=0}^{m}[0 \le k \le 1]\binom{m}{k}g(l-1, m-k)
          \]
          becomes
          \[
            g(l, m) = g(l-1, m) + mg(l-1, m-1)
          \]
    \item Identify a sufficient set of base cases
          \begin{itemize}
            \item e.g., \structure{$\{\, g(0, m), g(l, 0) \,\}$}
          \end{itemize}
  \end{enumerate}
\end{frame}

% TODO:
% 2) maybe space out
% 3) add labels for each step
% 4) add question marks that then change to 0^m and 1
% 5) make arrows gray
% 6) make the last line (except C++ code) appear last
% 8) mark which steps will be explained in more detail
\begin{frame}[fragile]{Knowledge Compilation Workflow (2/2)}
  \centering For each base case:

  \begin{tikzcd}
    g(0, m) \arrow[d] & \phantom{-} \arrow[d, phantom] & g(l, 0) \arrow[d] \\
    \forall x \in \Gamma\text{. }S(x) \lor \neg S(x) \arrow[d] & \phantom{-} & \top \arrow[d] \\
    g(0, m) = 0^m \arrow[rd] & & g(l, 0) = 1 \arrow[ld] \\
    & \mathcal{E} \arrow[d] & \\
    \text{Domain sizes} \arrow[r] & \text{C++ code} \arrow[r] & \text{Model count}
  \end{tikzcd}
\end{frame}

>>>>

\documentclass{beamer}
\usepackage[utf8]{inputenc}
\usepackage[UKenglish]{babel}
\usepackage[UKenglish]{isodate}
\usepackage{tikz}
\usepackage[linesnumbered,ruled,vlined]{algorithm2e}
\usepackage{complexity}
\usepackage{fontawesome}
\usepackage{dashbox}

@@

\tikzset{
    invisible/.style={opacity=0},
    visible on/.style={alt={#1{}{invisible}}},
    alt/.code args={<#1>#2#3}{%
      \alt<#1>{\pgfkeysalso{#2}}{\pgfkeysalso{#3}}%
  }
}

@@

% \begin{frame}{Knowledge Compilation Workflow}
%   \centering
%   \begin{tikzpicture}
%     % Top row
%     \node[anchor=west] at (-0.5, 0) (formula) {$\phi$};
%     \node[draw,rounded rectangle] at (3, 0) (compilewithbasecases) {\CompileWithBaseCases};
%     \node[draw,rounded rectangle] at (8, 0) (compilation) {Compile to C++};

%     % Bottom row
%     \node[draw,rounded rectangle,dashed] at (8, -3) (cpp) {C++ code};
%     \node[anchor=west] at (-1, -3) (sizes) {Domain sizes};
%     \node at (8, -4) (count) {Model count};

%     % Modules
%     \node[draw,rounded rectangle,anchor=north west] at (1.2, -1) (crane) {\Compile};
%     \node[draw,rounded rectangle,right=0.1cm of crane] (findbasecases) {\FindBaseCases};
%     \node[draw,rounded rectangle,anchor=north west] at (1.2, -1.6) (propagate) {\Propagate};
%     \node[draw,rounded rectangle,right=0.1cm of propagate] (simplify) {\Simplify};

%     % Bounding box and its name
%     \begin{scope}[on background layer]
%       \node[draw,fit={(compilewithbasecases) (compilation) (crane) (findbasecases) (propagate)},inner ysep=7pt,yshift=5pt,fill=gray!10] {};
%     \end{scope}
%     \node at (1.5, 0.5) {\Cranetwo};

%     % Brace and its arrow
%     \node[fit=(crane)(findbasecases)(propagate)(simplify)] (uses) {};
%     \draw[decorate,decoration={brace}] (uses.north west) -- (uses.north east) node[midway] (brace) {};
%     \draw[-Latex,dashed] (compilewithbasecases) -- node[midway,left] {uses} (brace);

%     % All other arrows
%     \draw[-Latex] (formula) -- (compilewithbasecases);
%     \draw[-Latex] (compilewithbasecases) -- node[above] {$\mathcal{E}$} (compilation);
%     \draw[-Latex] (compilation) -- (cpp);
%     \draw[-Latex] (sizes) -- (cpp);
%     \draw[-Latex] (cpp) -- (count);
%   \end{tikzpicture}
% \end{frame}

% TODO: 5 slides explaining the solution (maybe including the theorems)
% 0. Preprocess
% 1. Use Crane to compile into functions
% 2. Simplify
% 3. FindBaseCases TODO
% 4. For each base case
% 4.1. For each constant (TODO?)
% 4.1.1. Propagate TODO
% 5. Compile to C++

% TODO: (here and on the next slide) mark which steps will be explained in more detail
\begin{frame}{Knowledge Compilation Workflow (1/2)}
  \begin{enumerate}
    \item Use \textsc{Crane} to compile \structure{$\phi$} into a set of
          equations \structure{$\mathcal{E}$} \pause
    \item Simplify them, e.g.,
          \[
            g(l, m) = \sum_{k=0}^{m}[0 \le k \le 1]\binom{m}{k}g(l-1, m-k)
          \]
          becomes
          \[
            g(l, m) = g(l-1, m) + mg(l-1, m-1)
          \] \pause
    \item Identify a sufficient set of base cases
          \begin{itemize}
            \item e.g., \structure{$\{\, g(0, m), g(l, 0) \,\}$}
          \end{itemize}
  \end{enumerate}
\end{frame}

\begin{frame}[fragile]{Knowledge Compilation Workflow (2/2)}
  \centering \structure{4.} For each base case:

  \begin{tikzcd}
    g(0, m) \arrow[d, gray, visible on=<2->] & \phantom{-} \arrow[d, phantom, visible on=<2->, "\text{\structure{4.1.} Construct the corr.\ sentence}"] & g(l, 0) \arrow[d, gray, visible on=<2->] \\
    |[visible on=<2->]| \forall x \in \Gamma\text{. }S(x) \lor \neg S(x) \arrow[d, gray, visible on=<3->] & \phantom{-} \arrow[d, phantom, visible on=<3->, "\text{\structure{4.2.} Recurse}"] & |[visible on=<2->]| \top \arrow[d, gray, visible on=<3->] \\
    |[visible on=<3->]| g(0, m) = \alt<4->{0^m}{\text{???}} \arrow[rd, gray, visible on=<5->] & |[visible on=<5->]| \text{\structure{4.3.} Add to }\structure{\mathcal{E}} & |[visible on=<3->]| g(l, 0) = \alt<4->{1}{???} \arrow[ld, gray, visible on=<5->] \\
    & |[visible on=<5->]| \mathcal{E} \arrow[d, gray, visible on=<6->, "\text{\color{black}{\structure{5.} Compile to C++}}"] & \\
    |[visible on=<7->]| \text{Domain sizes} \arrow[r, gray, visible on=<7->] & |[visible on=<6->]| \text{\dbox{C++ code}} \arrow[r, gray, visible on=<7->] & |[visible on=<7->]| \text{Model count}
  \end{tikzcd}
\end{frame}
